---
title: "Under Construction"
author: "Jens von Bergmann"
date: '2017-08-29'
output:
  html_document: default
  html_notebook: default
vignette: >
  %\VignetteIndexEntry{Under Construction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette demonstrates how to pull in under constructino data from CMHC using the `cmhc` package,
link it with geographic data from [CensusMapper](https://censusmapper.ca) using the `cancensus`
package and map the under construction data.
## CMHC data

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(cmhc)

cma="Toronto"
year=2017
month=7

# properly parse integer with thousands separator. Yes CMHC sends down formatted number strings.
parse_integer <- function(x){
  return(as.integer(sub(",", "", x, fixed = TRUE)))
}

census_cma=census_geography_list[[cma]]
cma_header=substr(census_cma, nchar(census_cma)-2,nchar(census_cma))

#get all under construction data for Vancouver and pad CT GeoUIDs.
under_construction <- get_cmhc(cmhc_completion_params(
  geography_id = cmhc_geography_list[[cma]],
  year = year,
  month = month))
as_integer=c("Single","Semi-Detached","Row","Apartment","All")
under_construction <- under_construction %>%
  mutate(GeoUID = paste0(cma_header,X1)) %>%
  mutate_at(intersect(names(under_construction), as_integer), funs(parse_integer))

```

## Geographic data
```{r, message=FALSE, warning=FALSE}
library(cancensus)
#options(cancensus.api_key='your API key')
geos <- get_census(dataset = 'CA16', regions=list(CMA=census_cma),level='CT',geo_format='sf')
```

## Joining the data
```{r}
geos <- inner_join(geos,under_construction, by="GeoUID")
```


## Graph
```{r}
bg_color="#c0c0c0"
theme_opts<-list(theme(panel.grid.minor = element_blank(),
                       #panel.grid.major = element_blank(), #bug, not working
                       panel.grid.major = element_line(colour = bg_color),
                       panel.background = element_rect(fill = bg_color, colour = NA),
                       plot.background = element_rect(fill=bg_color, size=1,linetype="solid"),
                       axis.line = element_blank(),
                       axis.text.x = element_blank(),
                       axis.text.y = element_blank(),
                       axis.ticks = element_blank(),
                       axis.title.x = element_blank(),
                       axis.title.y = element_blank()))

```

After defining a basic theme we can go ahead and map the data.
```{r, fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
library(ggplot2)
library(sf)
library(RColorBrewer)

breaks=c(-Inf,1,25,50,100,250,500,750,1000,2000,Inf)
labels <- c(paste0("0 - ",breaks[2]))
for(i in 2:(length(breaks)-2)){
  labels[i] = paste(breaks[i],breaks[i+1], sep=" - ")
}
labels[length(breaks)-1]=paste0("Over ",breaks[length(breaks)-1])
#colors=c("darkred",(RColorBrewer::brewer.pal(length(labels)-1,"YlGnBu")))
colors=RColorBrewer::brewer.pal(length(relative_labels),"PiYG")

# categorize the numbers under contruction
geos$categories <- geos$All %>% cut(breaks=breaks, labels=labels)
total=sum(geos$All)

ggplot(geos) +
  geom_sf(aes(fill = categories), size = 0.05) +
  scale_fill_manual(labels=labels, values=colors, name = "# Units") +
  ggtitle(paste0(cma, " CMA Under Construction ",month,"-",year," (",prettyNum(total,big.mark = ",")," total)")) +
  theme_opts

ggsave('../images/under_construction.png',width=10,height=10)

```

## Mixing with Census Data
We can also mix in census data. For example we can show the construction activity relative to the 2016 dwelling stock. 

```{r, fig.height=5, fig.width=5, message=FALSE, warning=FALSE}

geos$relative_activity <- geos$All / geos$Dwellings 
total=sum(geos$All)/sum(geos$Dwellings)
relative_breaks <- c(-Inf,0.001,0.0025,0.005,0.0075,0.01,0.015,0.025,0.05,0.1,0.5,Inf)
geos$relative_categories <- geos$relative_activity %>% cut(breaks=relative_breaks)

relative_labels <- c(paste0("0 - ",prettyNum(relative_breaks[2]*10000,big.mark = ",")))
for(i in 2:(length(relative_breaks)-2)){
  relative_labels[i] = paste0(prettyNum(relative_breaks[i]*10000,big.mark = ","), " - ",prettyNum(relative_breaks[i+1]*10000,big.mark = ","))
}
relative_labels[length(relative_breaks)-1]=paste0("Over ",prettyNum(relative_breaks[length(relative_breaks)-1]*10000,big.mark = ","))
relative_colors=RColorBrewer::brewer.pal(length(relative_labels),"RdYlGn")
title=paste0(cma, " CMA Under Construction ",month,"-",year,
             " per 10,000 existing dwellings  (",
             prettyNum(round(total*10000),big.mark = ","),
             " on average)")

ggplot(geos) +
  geom_sf(aes(fill = relative_categories), size = 0.05) +
  scale_fill_manual(labels=relative_labels, values=relative_colors, name = "# Units per 10k\nExisting 2016",na.value="#808080") +
  ggtitle(title) +
  theme_opts
```

We could also just as easily choose other census variables to try and understand why dwelling unit construction is higher in some areas than in others. For example.

