---
title: "Under Construction"
author: "Jens von Bergmann"
date: '2017-08-29'
output:
  html_document: default
  html_notebook: default
vignette: >
  %\VignetteIndexEntry{Under Construction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette demonstrates how to pull in under constructino data from CMHC using the `cmhc` package,
link it with geographic data from [CensusMapper](https://censusmapper.ca) using the `cancensus`
package and map the under construction data.
## CMHC data

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(cmhc)

cma="Toronto"
year=2017
month=7

# properly parse integer with thousands separator. Yes CMHC sends down formatted number strings.
parse_integer <- function(x){
  return(as.integer(sub(",", "", x, fixed = TRUE)))
}

census_cma=census_geography_list[[cma]]
cma_header=substr(census_cma, nchar(census_cma)-2,nchar(census_cma))

#get all under construction data for Vancouver and pad CT GeoUIDs.
under_construction <- get_cmhc(cmhc_completion_params(
  geography_id = cmhc_geography_list[[cma]],
  year = year,
  month = month))
as_integer=c("Single","Semi-Detached","Row","Apartment","All")
under_construction <- under_construction %>%
  mutate(GeoUID = paste0(cma_header,X1)) %>%
  mutate_at(intersect(names(under_construction), as_integer), funs(parse_integer))

```

## Geographic data
```{r, message=FALSE, warning=FALSE}
library(cancensus)
#options(cancensus.api_key='your API key')
geos <- get_census(dataset = 'CA16', regions=list(CMA=census_cma),level='CT',geo_format='sf')
```

## Joining the data
```{r}
geos <- inner_join(geos,under_construction, by="GeoUID")
```


## Graph
```{r}
bg_color="#c0c0c0"
theme_opts<-list(theme(panel.grid.minor = element_blank(),
                       #panel.grid.major = element_blank(), #bug, not working
                       panel.grid.major = element_line(colour = bg_color),
                       panel.background = element_rect(fill = bg_color, colour = NA),
                       plot.background = element_rect(fill=bg_color, size=1,linetype="solid"),
                       axis.line = element_blank(),
                       axis.text.x = element_blank(),
                       axis.text.y = element_blank(),
                       axis.ticks = element_blank(),
                       axis.title.x = element_blank(),
                       axis.title.y = element_blank()))

```

After defining a basic theme we can go ahead and map the data.
```{r, fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
library(ggplot2)
library(sf)
library(RColorBrewer)

breaks=c(-Inf,1,25,50,100,250,500,750,1000,2000,Inf)
labels <- c(paste0("0 - ",breaks[2]))
for(i in 2:(length(breaks)-2)){
  labels[i] = paste(breaks[i],breaks[i+1], sep=" - ")
}
labels[length(breaks)-1]=paste0("Over ",breaks[length(breaks)-1])
#colors=c("darkred",(RColorBrewer::brewer.pal(length(labels)-1,"YlGnBu")))
labels=factor(labels, levels=labels)
colors=setNames(RColorBrewer::brewer.pal(length(labels),"PiYG"),labels)
#colors=factor(as.character(colors),levels=as.character(colors))

# categorize the numbers under contruction
geos$categories <- geos$All %>% cut(breaks=breaks, labels=labels)
total=sum(geos$All)

ggplot(geos) +
  geom_sf(aes(fill = categories), size = 0.05) +
  scale_fill_manual(labels=labels, values=colors, name = "# Units") +
  ggtitle(paste0(cma, " CMA Under Construction ",month,"-",year," (",prettyNum(total,big.mark = ",")," total)")) +
  theme_opts

ggsave('../images/under_construction.png',width=10,height=10)

```

## Mixing with Census Data
We can also mix in census data. For example we can show the construction activity relative to the 2016 dwelling stock. 

```{r, fig.height=5, fig.width=5, message=FALSE, warning=FALSE}

geos$relative_activity <- geos$All / geos$Dwellings 
relative_total=sum(geos$All)/sum(geos$Dwellings)
relative_breaks <- c(-Inf,0.001,0.0025,0.005,0.0075,0.01,0.015,0.025,0.05,0.1,0.5,Inf)

relative_labels <- c(paste0("0 - ",prettyNum(relative_breaks[2]*10000,big.mark = ",")))
for(i in 2:(length(relative_breaks)-2)){
  relative_labels[i] = paste0(prettyNum(relative_breaks[i]*10000,big.mark = ","), " - ",prettyNum(relative_breaks[i+1]*10000,big.mark = ","))
}
relative_labels[length(relative_breaks)-1]=paste0("Over ",prettyNum(relative_breaks[length(relative_breaks)-1]*10000,big.mark = ","))
relative_colors=setNames(RColorBrewer::brewer.pal(length(relative_labels),"RdYlGn"),relative_labels)

geos$relative_categories <- geos$relative_activity %>% cut(breaks=relative_breaks, labels=relative_labels)

title=paste0(cma, " CMA Under Construction ",month,"-",year,
             " per 10,000 existing dwellings  (",
             prettyNum(round(relative_total*10000),big.mark = ","),
             " on average)")

ggplot(geos) +
  geom_sf(aes(fill = relative_categories), size = 0.05) +
  scale_fill_manual(labels=relative_labels, values=relative_colors, name = "# Units per 10k\nExisting 2016",na.value="#808080") +
  ggtitle(title) +
  theme_opts
```

We could also just as easily choose other census variables to try and understand why dwelling unit construction is higher in some areas than in others. For example.

## Annex
Everyone seems to be talking about this area called the Annex these days. Let's check what's under development there right now,
so let's zoom in a bit and overlay the neighbourhood boundary. While most of the Annex has essentially zero development happening right now, the north-eastern corner that also holds the subject site that irks everyone's interest these days does seems to see some moderate development activity. It is also evident that the development activity is spread very unevenly, and some areas
are recieving an order of magnitude more active development.


```{r, fig.height=5, fig.width=5, message=FALSE, warning=FALSE}

annex <- read_sf('../data/annex.geojson')
site <- data.frame(long=-79.40125107,lat=43.674595)

ggplot() +
  geom_sf(data=geos, aes(fill = categories), size = 0.05) +
  ggtitle(paste0(cma, " Under Construction ",month,"-",year)) +
  scale_fill_manual(values=colors,  breaks =labels, labels=labels, name = "# Units") +
  theme_opts +   
  geom_point(data=site,aes(long,lat),colour="blue",size=2,show_guide = FALSE) +
  geom_sf(data=annex, aes(fill=NA),color="blue",show_guide = FALSE) +
  scale_color_discrete(guide=FALSE) +
  #guides(fill=FALSE) +
  coord_sf(xlim=c(-79.3,-79.5), ylim=c(43.6,43.75))  

ggsave('../images/under_construction_annex.png',width=10,height=10)

```
Lastly, let's check exactly how many units are under construction in the Annex. The Annex is covered by 3 CTs, so this is
quite easy. Some of the construction might be outside of the Annex, the sliver on the north side of Dupond Street and south of the railroad tracks. But this gives a simple upper bound.

```{r}
annex_cts <- c("5350092.00","5350091.01","5350091.02")
paste0(sum(geos$All[geos$GeoUID %in% annex_cts]), " units under construction in the Annex")

```

