---
title: "Vacancy Rates and Rent Changes"
author: "Jens von Bergmann"
date: '2017-08-29'
output:
  html_document: default
  html_notebook: default
vignette: |
  %\VignetteIndexEntry{Vacancy Rates and Rent Changes} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}
---

This vignette uses the `cmhc` package to download vacancy and rent change data for several areas and visualize them on the same graph to highlight the relationship between the two.

To install and use `cmhc` simply download the repository from Github. 
```{r, message=FALSE, warning=FALSE}
#devtools::install_github("mountainmath/cmhc")
library(cmhc)
```


## Getting the Data

First we read in the data using the `cmhc` package and join the tables we want and tidy up.`cmhc` is preloaded with a list of pre-loaded parameters, `cmhc_geography_list` to access CMHC data for the following cities:
```{r, message=FALSE, warning=FALSE}
cmhc_geography_list
```

The CMHC API is a bit of a mess. `cmhc` uses several internal functions to access data via the CMHC API. The function below makes calls to the CMHC API and returns vacancy and rent price data for a given CMHC city id in a tidy way that we can then use for analysis or graphing. 

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)

# A function to read and tidy CMHC data
vacancy_rent_table_for <- function(geography,geography_type="CMA"){
  region_params=cmhc_region_params(geography = geography,type=geography_type)
  params=cmhc_timeseries_params(table_id = cmhc_table_list["Rms Vacancy Rate Time Seris"],region=region_params)
  dat_vacancy=get_cmhc(params)
  title_x=attr(dat_vacancy,"title")
  dat_vacancy <- dat_vacancy %>% 
    select("X1","Total") %>%
    #mutate(Total=as.numeric(as.character(Total))/100) %>%
    rename(vacancy_rate=Total, X=X1)
  params=cmhc_timeseries_params(table_id = cmhc_table_list["Rms Rent Change Time Seris"],region=region_params)
  dat_rent_change=get_cmhc(params)
  title_y=attr(dat_rent_change,"title")
  dat_rent_change <- dat_rent_change %>%
    select("X1","Total") %>%
    #mutate(Total=as.numeric(as.character(Total))/100) %>%
    rename(rent_change=Total, X=X1)
  dat=inner_join(dat_vacancy,dat_rent_change,by="X") %>% rename(Year=X)
  attr(dat,"region")=paste0(geography," ",geography_type)
  attr(dat,"labels")=c(title_x,title_y)
  return(dat)
}
```

Let's take a look at this data now.

```{r, message=FALSE, warning=FALSE}
# cmhc_geography_list
cgy <- vacancy_rent_table_for("Calgary")
head(cgy)
tor <- vacancy_rent_table_for("Toronto")
head(tor)
van <- vacancy_rent_table_for("Vancouver")
head(van)
```

And combine it into a single data frame for comparing directly:

```{r, message=FALSE, warning=FALSE}
cmhc <- bind_rows(cgy, tor, van, .id = "city")

# Rename id var
cmhc <- cmhc %>% 
  # Gather into tidy long format
  gather(key = "Series", value = "Rate",vacancy_rate:rent_change) %>%
  # Tidy up the names and variables
  mutate(city = case_when(
    .$city == 1 ~ "Calgary",
    .$city == 2 ~ "Toronto",
    .$city == 3 ~ "Vancouver"),
    Series = case_when(
    .$Series == "vacancy_rate" ~ "Vacancy Rate",
    .$Series == "rent_change" ~ "Rent Change"),
    Year = as.numeric(substr(Year,1,4)))
```

## Plot the data

With the data all tidy, we can now plot it easily. 

```{r, message=FALSE, warning=FALSE}
library(ggplot2)

plot_comparison <- ggplot(cmhc, aes(x = Year, y = Rate, color = Series)) +
  geom_line() + 
  geom_point() +
  facet_grid(city~.) + 
  labs(title="Vacancy Rate vs Change in Rent", 
       subtitle ="Select Cities",
       caption="Source: CMHC Rms, code: https://github.com/mountainMath/cmhc") +
  scale_y_continuous(labels = scales::percent) +
  xlab("") +
  scale_color_manual(labels = c("% Rent Change\n(fixed sample)","Vacancy Rate"), values = c("darkgreen", "steelblue"), name = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
plot_comparison

ggsave('../images/vacancy_rent_change.png')

```



We can narrow things down by city instead of CMA.

cmhc_geography_csd_list


```{r, message=FALSE, warning=FALSE}
# 
cgy <- vacancy_rent_table_for("Calgary","CSD")
tor <- vacancy_rent_table_for("Toronto","CSD")
van <- vacancy_rent_table_for("Vancouver","CSD")
cmhc <- bind_rows(cgy, tor, van, .id = "city")
# Rename id var
cmhc <- cmhc %>% 
  # Gather into tidy long format
  gather(key = "Series", value = "Rate",vacancy_rate:rent_change) %>%
  # Tidy up the names and variables
  mutate(city = case_when(
    .$city == 1 ~ "Calgary CSD",
    .$city == 2 ~ "Toronto CSD",
    .$city == 3 ~ "Vancouver CSD"),
    Series = case_when(
    .$Series == "vacancy_rate" ~ "Vacancy Rate",
    .$Series == "rent_change" ~ "Rent Change"),
    Year = as.numeric(substr(Year,1,4)))
plot_comparison <- ggplot(cmhc, aes(x = Year, y = Rate, color = Series)) +
  geom_line() + 
  geom_point() +
  facet_grid(city~.) + 
  labs(title="Vacancy Rate vs Change in Rent", 
       subtitle ="Select Cities",
       caption="Source: CMHC Rms, code: https://github.com/mountainMath/cmhc") +
  scale_y_continuous(labels = scales::percent) +
  xlab("") +
  scale_color_manual(labels = c("% Rent Change\n(fixed sample)","Vacancy Rate"), values = c("darkgreen", "steelblue"), name = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
plot_comparison

ggsave('../images/vacancy_rent_change.png')

```
