---
title: "Vacancy Rates and Rent Changes"
author: "Jens von Bergmann"
date: "2017-08-29"
output:
  html_document: default
  html_notebook: default
vignette: >
  %\VignetteIndexEntry{Vacancy Rates and Rent Changes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette uses the `cmhc` package to download vacancy and rent change data for several areas and visualize them on the same graph to highlight the relationship between the two.

## Getting the Data
First we read in the data using the `cmhc` package and join the tables we want and tidy up.

```{r, message=FALSE, warning=FALSE}
library(cmhc)
library(dplyr)
library(ggplot2)
library(tidyr)

# read in the data and tidy up
vacancy_rent_table_for <- function(geography_id){
  dat_vacancy=get_cmhc(cmhc_vacancy_history_params(geography_id))
  region=attr(dat_vacancy,"region")
  title_x=attr(dat_vacancy,"title")
  dat_vacancy <- dat_vacancy %>% 
    select("X","Total") %>%
    mutate(Total=as.numeric(as.character(Total))/100) %>%
    rename(vacancy_rate=Total)
  dat_rent_change=get_cmhc(cmhc_rent_change_history_params(geography_id)) 
  title_y=attr(dat_rent_change,"title")
  dat_rent_change <- dat_rent_change %>%
    select("X","Total") %>%
    mutate(Total=as.numeric(as.character(Total))/100) %>%
    rename(rent_change=Total)
  dat=inner_join(dat_vacancy,dat_rent_change,by="X") %>% rename(Year=X)
  attr(dat,"region")=region
  attr(dat,"labels")=c(title_x,title_y)
  return(dat)
}
```

## Plot the data
With an easy way to get the data, we now define the plot

```{r, message=FALSE, warning=FALSE}
plot_geography <- function(data) {
  plot_data <- data %>% 
    gather(key="series",value="Rate", "vacancy_rate", "rent_change") %>% 
    mutate(Rate=as.numeric(as.character(Rate))) 

  ggplot(data=plot_data, aes(x = Year, y = Rate, group=series, color=series)) +
  geom_line() +
  geom_point() + 
  labs(title=paste0(attr(data,"region")," CMA"), caption="Source: CMHC Rms, code: https://github.com/mountainMath/cmhc") +
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(labels = c("% Rent Change\n(fixed sample)","Vacancy Rate"), values = c("darkgreen", "steelblue")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
}

```

Equipped with these two functions we can now plot the data for various census metropolitan areas.

```{r, message=FALSE, warning=FALSE, fig.height=3, fig.width=5}
plot_geography(vacancy_rent_table_for(cmhc_geography_list$Vancouver))
plot_geography(vacancy_rent_table_for(cmhc_geography_list$Calgary))
plot_geography(vacancy_rent_table_for(cmhc_geography_list$Toronto))

```

